---
title: "Homework 2"
author: "Jerry Chao"
date: "9/24/2020"
output: github_document
---

## This is my solution to Homework 2

## Problem 0
# I have created a GitHub repo as well as local R Project named "p8105_hw2_jyc2171", followed by one single .Rmd file named "p8105_hw2_jyc2171.Rmd" that renders to GitHub repo.  A subdirectory was created to use the data files Mr. Trash Wheel (Problem 1), NYC transit data (Problem 2), and FiveThiryEight (Problem 3) using relative paths.  This homework shall be submitted via courseworks as a URL to my GitHub repo.  Thanks.

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1 - Mr. Trash Wheel dataset.  Read and clean dataset.


```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data! for 2018 and 2017.

```{r}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1 
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
    
precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1 
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)
  
left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, MD.  As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster.   The dataset contains information on year, month, and trash collected, including some specific kinds of trash.  There are a total of `r nrow(trashwheel_df)` rows in our final dataset.  Additional data sheets include monthly precipitation data.  

